---
title: "materialHardship"
author: "Adaeze Ajoku"
date: "5/26/2017"
output: html_document
---

# Step 0- Prepare for data analysis

## Load in libraries
```{r}
library(readr)
library(randomForest)
library(mice)
library(dplyr)
library(glmnet)
```

## Load in data
```{r}
background <- read.csv("~/Desktop/Fragile Families/background.csv", header=T)
train <- read.csv("~/Desktop/Fragile Families/train.csv", header=T)
background.tr <- semi_join(background, train)
```

# Approach 1~ Select what I believe are important variables

## Year 0
```{r}
year0 <- c('m1a4', 'm1a11a', 'm1a11b', 'm1a11c', 'm1a11d', 'm1a13', 'm1a15', 'cm1age', 'cm11bw', 'm1b6a', 'm1b9b3', 'm1b10a3', 'm1b12a', 'm1b20', 'cm1cohf', 'm1d1c', 'm1e4a', 'm1f3', 'm1f4', 'cm1edu', 'm1j1b', 'm1j1e', 'm1j2a', 'm1j2b', 'm1j3', 'cm1adult', 'cm1kids', 'f1a6', 'f1a6a', 'f1b6a', 'f1b9b3', 'f1b10b3', 'f1d1e', 'f1e3', 'f1e7a', 'f1f3', 'f1f4', 'f1f8', 'cf1edu', 'f1k1a', 'f1k1c', 'cf1adult', 'cf1kids')
```

## Year 1
```{r}
year1 <- c()
```

## Year 3
## Year 5
## Year 9

## Final Dataset
```{r}
fin <- background[,c('challengeID', year0)]
```
# Approach 2~ Dimension reduction

## LASSO
Turns out LASSO cannot handle missing (NA) values.

Missing Data Interpretation:
-9 *Not in wave* – Did not participate in survey/data collection component
-8 Out of range – Response not possible; rarely used
-7 Not applicable (also -10/-14) – Rarely used for survey questions
-6 *Valid skip* – Intentionally not asked question; question does not apply to respondent or response known based on prior information.
-5 Not asked “Invalid skip” – Respondent not asked question in the version of the survey they received.
-3 Missing – Data is missing due to some other reason; rarely used
-2 *Don’t know* – Respondent asked question; Responded “Don’t Know”.
-1 *Refuse* – Respondent asked question; Refused to answer question
```{r complete cases}
# (a) What if I only use the rows with complete cases (ie no NAs)?
btr.all <- data.frame(background.tr, materialHardship = train$materialHardship)
btr.cc <- complete.cases(btr.all) #all False
#There aren't any rows without at least one NA. So this won't work.
```

```{r multiple imputation prep}
# (b) How about doing multiple imputation?

## (i) Set negative values to NA
negs <- c(-1, -2, -3, -5, -6, -7, -8, -9, -10, -14)
neg2NA <- function(x){
  ifelse(x %in% negs, NA, x)
}

background.tr <- as.data.frame(sapply(background.tr, function(x) neg2NA(x)))

## (ii) Some columns are entirely NA or more than 50% NA. These rows I'd like to eliminate from consideration.
table(background.tr$m1lenhr, useNA='ifany') #for example

miss_pat <- md.pattern(background.tr) 
#There doesn't seem to be any pattern to the missing data

#I cannot use aggr() in the VIM package because there are too many variables to graphically represent the missing data pattern

miss_pairs <- md.pairs(background.tr) #takes too long to compute

na_percent <- function(x){
  sum(is.na(x))/nrow(background.tr)*100
}

background.tr_na <- background.tr[,which(sapply(background.tr, function(x) na_percent(x)) > 50)]

background.tr_cl <- background.tr[,!(names(background.tr) %in% names(background.tr_na))]

# Now that I have a more manageable dataset I can look towards multiple imputation. However I must specify a mice method for each column. 

## (ii) For the columns with no missing values, I will set the imputation method to "". If it is a numeric I will use method "norm", if is a 2-level factor I will use method "logreg", and if it is a 2+ level factor I will use "polyreg". 

background.tr_cl.na <- background.tr_cl[,which(sapply(background.tr_cl, function(x) sum(is.na(x))) != 0)]

#temp <- background.tr_cl.na[,!(names(background.tr_cl.na) %in% norm1 | names(background.tr_cl.na) %in% logreg1 | names(background.tr_cl.na) %in% polyreg1)]
temp2 <- background.tr_cl.na[,!(names(background.tr_cl.na) %in% names(temp))]


for(i in names(temp2)){
  if(i %in% polyreg1){
    print(i)
  }
}
```

```{r assigning method type}
# norm for numeric types
norm1 <- c("m1citywt", "m1e1d8", "m1citywt_rep1", "m1citywt_rep2", "m1citywt_rep3", "m1citywt_rep4",  "m1citywt_rep5", "m1citywt_rep6", "m1citywt_rep7", "m1citywt_rep8", "m1citywt_rep9", "m1citywt_rep10", "f1citywt", "f1j15a", "f1citywt_rep1", "f1citywt_rep2", "f1citywt_rep3", "f1citywt_rep4", "f1citywt_rep5", "f1citywt_rep6", "f1citywt_rep7", "f1citywt_rep8", "f1citywt_rep9", "f1citywt_rep10", "m2citywt", "m2citywt_rep1", "m2citywt_rep2", "m2citywt_rep3", "m2citywt_rep4", "m2citywt_rep5", "m2citywt_rep6", "m2citywt_rep7", "m2citywt_rep8", "m2citywt_rep9", "m2citywt_rep10", "f2citywt", "f2citywt_rep1", "f2citywt_rep2", "f2citywt_rep3", "f2citywt_rep4", "f2citywt_rep5", "f2citywt_rep6", "f2citywt_rep7", "f2citywt_rep8", "f2citywt_rep9", "f2citywt_rep10", "m3citywt", "m3citywt_rep1", "m3citywt_rep2", "m3citywt_rep3", "m3citywt_rep4", "m3citywt_rep5", "m3citywt_rep6", "m3citywt_rep7", "m3citywt_rep8", "m3citywt_rep9", "m3citywt_rep10", "f3citywt", "f3citywt_rep1", "f3citywt_rep2", "f3citywt_rep3", "f3citywt_rep4", "f3citywt_rep5", "f3citywt_rep6", "f3citywt_rep7", "f3citywt_rep8", "f3citywt_rep9", "f3citywt_rep10", "m4citywt", "m4citywt_rep1", "m4citywt_rep2", "m4citywt_rep3", "m4citywt_rep4", "m4citywt_rep5", "m4citywt_rep6", "m4citywt_rep7", "m4citywt_rep8", "m4citywt_rep9", "m4citywt_rep10", "m2c37a4", "m2g5b_a1", "f2fh23a", "f2k15a", "f4k18", "f4k26c", "f4l6b", "f4l9b", "m4j13a", "hv4a12", "hv4a14",  "hv4b1", "hv4b2", "hv4b3", "hv4b4", "hv4b5", "hv4b23", "hv4b24",  "hv4e3", "hv4e4", "hv4e5", "hv4food_exp", "hv4accdt")

# logreg for 2 factors
logreg1 <-  c("cm1finjail", "cmf1finjail", "cf1natsm", "cf1citsm", "cf1finjail", "cm2natsm", "cm2natsmx", "cm2citsm", "cm2finjail", "cm2fevjail", "cf2natsm", "cf2natsmx", "cf2citsm", "cf2finjail", "cmf2finjail", "cf2fevjail", "cmf2fevjail", "cc4natsm", "cc4natsmx", "cc4citsm", "cmf4finjail", "cmf4fevjail", "f4l6", "cc2natsm", "cc2natsmx", "cc2citsm", "cm3natsm", "cm3natsmx", "cm3citsm", "cm3finjail", "cm3fevjail", "m3l9", "cf3natsm", "cf3natsmx", "cf3citsm", "cf3finjail", "cmf3finjail", "cf3fevjail", "cmf3fevjail", "cc3natsm", "cc3natsmx", "cc3citsm", 'cm4natsm', "cm4natsmx", "cm4citsm", "cm4finjail", "cm4fevjail", "m4k18", "m4k24c", "m4k26c", "m4l9b", "hv4a7", "hv4a9", "hv4a16", "hv4a20", "hv4a22", "hv4a29", "hv4b7", "hv4b9",  "hv4b11", "hv4b15", "hv4b17", "hv4b18", "hv4c2", "hv4c3", "hv4c5", "hv4c6", "hv4c9", "hv4c13", "hv4c14", "hv4d2", "hv4g15", "hv4g20", "hv4j1", "hv4sex_child", "hv4k0a",  "hv4k1", "hv4k3", "hv4k9")  

# polyreg for 2+ factors 
polyreg1 <- c("m1b9b11", "m1b10a11", "m1f7", "m1h3", "m1i11", "f1b9a11", "f1b10b11", "f1f7",  "m2a7c",  "m2b15a8", "m2b30g", "m2b31cp", "m2b32p", "m2c14", "m2c20p", "m2c27cp", "m2d3b7", "m2f2d3", "m2f2d4", "m2g8a", "m2k6b", "m2l9", "f2a4a", "f2b14a8", "f2b22", "f2fc19bp", "f2fc19c5", "f2g5b", "f2h12a6", "f2j4b6", "f2k3k", "m5a6f03", "m5d7", "m5f14b", "f5a4c", "f5a6f02", "f5a6f03", "f5b10x", "f5e1b", "f5f14b", "f5i7", "f5j5d", "p5l5b", "hv5_dspr", "f4a10h1", "f4c10a1", "f4c15a_4", "f4c21d_9", "f4c36", "f4c37a_8", "f4h1b", "f4i2", "f4j2a_8", "f4l5d", "f2k23f1d", "m3b3f", "m3c15a_4", "m3c38d_9", "f3b15_7", "f3c38c_9", "f3c41", "f3i2",  "m4a3i", "m4c10a2", "m4c21d_9", "m4c36", "m4f3a1_5", "m4f3a2_9", "m4h1b", "m4i14b", "m4i20b", "m4j2a_8", "m4j2b1_9", "m4l5c", "m4l5d", "hv4a1", "hv4a6", "hv4a8", "hv4a11", "hv4a21d1_ot", "hv4a21d2_ot", "hv4a21d3_ot", "hv4a27", "hv4a30", "hv4b8", "hv4b14", "hv4b14_6ot", "hv4b14_7ot", "hv4c1a", "hv4c1b", "hv4c1c", "hv4c1d", "hv4c1e", "hv4c1f",  "hv4c1g", "hv4c1h", "hv4c7", "hv4c8", "hv4c15", "hv4c16a", "hv4c16b", "hv4c16c", "hv4c17a", "hv4c17b", "hv4c17c", "hv4d1a", "hv4d1b", "hv4d1c", "hv4d1d", "hv4d1e", "hv4d14",  "hv4d15a", "hv4d15b",  "hv4d15c", "hv4d15d", "hv4d15e", "hv4d15f",  "hv4d15g",  "hv4d15h", "hv4d15i", "hv4d16", "hv4d17", "hv4d17_ot", "hv4e5_ot", "hv4f1a", "hv4f1b", "hv4f1c", "hv4f1d", "hv4f1e", "hv4f1f", "hv4f1g", "hv4f1h", "hv4f1j", "hv4f1k", "hv4f1l", "hv4f2a1", "hv4f2a1_ot1",  "hv4f2a1_ot2", "hv4f2b1", "hv4f2b1_ot1", "hv4f2b1_ot2", "hv4f3a",  "hv4f3b", "hv4f3c", "hv4f3d", "hv4f3e", "hv4g1", "hv4g2", "hv4g3", "hv4g4", "hv4g5", "hv4g6", "hv4g7", "hv4g8", "hv4g9", "hv4g10", "hv4g11", "hv4g12", "hv4g13", "hv4g14", "hv4g16", "hv4g17", "hv4g18", "hv4g19", "hv4g20_ot", "hv4g22_ot", "hv4h1", "hv4h2", "hv4h3", "hv4h4", "hv4h5", "hv4h6", "hv4h7", "hv4j5a_ot", "hv4j7_ot",  "hv4j11_ot", "hv4v8_ot", "hv4v9", "hv4v9_ot", "hv4k4per_ot", "hv4k5per_ot", "hv4k6", "hv4k6_ot", "hv4k7", "hv4k8", "hv4k8_ot", "hv4k10a_ot1", "hv4k10a_ot2", "hv4k11a_ot1", "hv4k11a_ot2", "hv4k12a_ot1", "hv4k13a_ot1",  "hv4k14a_ot1", "hv4k14a_ot2", "hv4l1", "hv4l2", "hv4l3", "hv4l4",  "hv4l5", "hv4l6", "hv4l7", "hv4l8", "hv4l9", "hv4l10", "hv4l11", "hv4l12", "hv4l14", "hv4l15", "hv4l16", "hv4l17", "hv4l18", "hv4l19", "hv4l20", "hv4l21", "hv4l22", "hv4l23", "hv4l24", "hv4l25", "hv4l26", "hv4l27", "hv4l28", "hv4l29", "hv4l30", "hv4l31",  "hv4l32", 'hv4l33', "hv4l34", "hv4l35", "hv4l36", "hv4l37", "hv4l38", "hv4l39", "hv4l40",  "hv4l41",  "hv4l42",  "hv4l43", "hv4l44", "hv4l45", "hv4l46", "hv4l47", "hv4l48", "hv4l49", "hv4l50", "hv4l51",  "hv4l52", "hv4l53", "hv4l54", "hv4l55", "hv4l56", "hv4l57", "hv4l58", "hv4l59", "hv4l60", "hv4l61", "hv4l62", "hv4l63", "hv4l64",  "hv4l65", "hv4l66", "hv4p6_ot", "hv4p8_ot1", "hv4r10a_ot2", "hv4s1_ot1", "hv4s1_ot2", "hv4s5_ot1_1", "hv4s5_ot1_2", "hv4s5_ot2_1")
```

```{r}
m.method <- rep("", length(background.tr_cl))

m.method <- ifelse(names(background.tr_cl) %in% norm1, "norm", ifelse(names(background.tr_cl) %in% logreg1, "logreg", ifelse(names(background.tr_cl) %in% polyreg1, "polyreg", m.method)))

## (iii) In determining the method for the other columns, if a column is a factor, I will make all negative values NA.



#The imputation process in R takes too long so I will read out my file to a csv and use fancy impute in Python
write.csv(background.tr_cl, "/Users/adaezeajoku/Desktop/Fragile\ Families/cleaned_background_tr.csv", row.names=F)




## (ii) Including the response variable in the imputation procedure
btr.mi <- mice(background.tr_cl[-1], method='norm')

## Doing separate imputation on background and on the response variable

## Doing separate imputation on background and using the mean for the NA values in the response

## Using the mean for NA values in background and the response variable

Can be either a single string, or a vector of strings with length ncol(data),
specifying the elementary imputation method to be used for each column in
data. If specified as a single string, the same method will be used for all columns.
The default imputation method (when no argument is specified) depends on the
measurement level of the target column and are specified by the defaultMethod
argument. Columns that need not be imputed have the empty method ''. See
details for more information.

lasso.fit <- glmnet(as.matrix(background.tr),train$materialHardship, family='gaussian', alpha = 1) 
```

## Principal Component Regression